Hydras \& Co.

### 2022-03-18 doc/gaia-chapter.tex
corrections chapter 5 (completed)

\subsection{Canonical sequences}
Canonical sequences are described in Chapter~\vref{chap:ketonen}, and implemented in
\href{../theories/html/hydras.Epsilon0.Canon.html}%
{\texttt{hydras.Epsilon0.Canon}}.
Module~\href{../theories/html/gaia_hydras.GCanon.html}%
{\texttt{gaia\_hydras.GCanon}}, imports definitions and results from that library.

@@

\subsection{Accessibility in \texorpdfstring{$\epsilon_0$}{epsilon\_0}}

@@

If $\beta$ is (strictly) accessible from $\alpha$, then
$\beta<\alpha$.

\inputsnippets{GPaths/pathToLT}

>>>>

\subsection{Canonical sequences}
\label{sect:gcanon}
Canonical sequences are described in Chapter~\vref{chap:ketonen}, and implemented in
\href{../theories/html/hydras.Epsilon0.Canon.html}%
{\texttt{hydras.Epsilon0.Canon}}.
Module~\href{../theories/html/gaia_hydras.GCanon.html}%
{\texttt{gaia\_hydras.GCanon}}, imports definitions and results from that library.

@@

\subsection{Accessibility in \texorpdfstring{$\epsilon_0$}{epsilon\_0}}
\label{sect:gpath}

@@

$\beta$ is (strictly) accessible from $\alpha$, if and only iff
$\beta<\alpha$.

\inputsnippets{GPaths/pathToLT}
\inputsnippets{GPaths/LTPathTo}

### 2022-03-18 doc/part-hydras.tex
corrections chapter 5 (completed)

\paragraph*{\gaiasign} Our library~\href{../theories/html/gaia_hydras.GCanon.html}{gaia\_hydras.GCanon} transposes the  function \texttt{canon} into \gaia's type \texttt{T1}.

In the present state of this library, the following specializations of \texttt{canon} are still used in some proofs or lemma statements.

@@

\section{Accessibility inside \texorpdfstring{$\epsilon_0$}{epsilon0} : paths}
\index{maths}{Ordinal numbers!Accessibility inside epsilon0}
\label{sect:pathes-intro}

@@

Let $s$ be the sequence $\langle s_0,s_1,\dots, s_{l-1} \rangle$. We describe the
existence of such a path with the notation $\alpha\xrightarrow [s]{}\beta$.

We say also that the considered path from $\alpha$ to $\beta$ \emph{starts at [index] $s_0$ and ends at $s_l$}.

@@

Note that, given $\alpha$ and $\beta$, where $\beta < \alpha$, the sequence $s$ which leads from $\alpha$ to $\beta$ is not unique.

Indeed, if $\alpha$ is a limit ordinal, the first element of $s$ can be any integer $i$ such that $\beta<\canonseq{\alpha}{i}$, and if $\alpha$ is a successor ordinal,
then the sequence $s$ can start with any positive integer.

@@

@@

@@

\pagebreak

@@

The sequence $s$, used to build the sequence of replication factors of the battle, depends on 
$\beta$, so we cannot be sure that the generated battle is a genuine standard battle.

The solution of this issue comes  once again from Ketonen and Solovay's article~\cite{KS81}. Instead of considering plain paths, i.e. sequences 
$\alpha_0=\alpha,\alpha_1,\dots,\alpha_k=\beta$ where $\alpha_{j+1}$ is equal
to $\canonseq{\alpha_j}{i_j}$ where $i_j$ is \emph{any} natural number, 
we consider various constraints on these sequences.
In particular, a path is called \emph{standard} if $i_{j+1} = i_j + 1$ for every $j<k$.
It  corresponds to a ``segment'' of some standard battles. 
Please note that the vocabulary on paths is ours, but all the concepts come really from~\cite{KS81}.

@@

@@

@@

Our proof follows the proof by Ketonen and Solovay, including its organization as a sequence of lemma.  Since it is a non-trivial proof, we will comment its main steps below.

@@

Applying \texttt{standard\_path\_to\_zero}, $0$ is reachable from $\alpha$ by some standard path  (see figure~\vref{fig:belle-preuve-1}).

@@

\subsection{Back to hydras}
\label{sec:standard-battles-cases}
We are now able to complete our proof that there exists no bounded variant for proving the termination of standard hydra battles. This proof can
be consulted in the module 
\href{../theories/html/hydras.Hydra.Epsilon0_Needed_Std.html}{Hydra.Epsilon0\_Needed\_Std}.
Please note that it has the same global structure as in section\ref{sec:free-battles-case} 
% ICI !
Applying the  lemmas  \texttt{Lemma2\_6\_1} of the module 
\href{../theories/html/hydras.Epsilon0.Paths.html\#Lemma2_6_1}%
{Epsilon0.pathS}   and 
\href{../theories/html/hydras.Epsilon0.Paths.html\#constant_to_standard_path}%
{\texttt{constant\_to\_standard\_path}},
we can convert any inequality $\beta<\alpha<\epsilon_0$ into a standard path from
$\alpha$ to  $\beta$, then into a fragment of a standard battle from 
$\iota(\alpha)$ to $\iota(\beta)$, hence the inequality $m(\iota(\beta))<m(\iota(\alpha))$.

>>>>

\paragraph*{\gaiasign} Our library~\href{../theories/html/gaia_hydras.GCanon.html}{gaia\_hydras.GCanon} transposes the  function \texttt{canon} into \gaia's type \texttt{T1}
(please see Sect.\~ref{sect:gcanon}).

\paragraph*{Remark}
In the present state of this library, the following specializations of \texttt{canon} are still used in some proofs or lemma statements. They are planned to be deprecated.

@@

\section{Accessibility inside \texorpdfstring{$\epsilon_0$}{epsilon0}\,: paths}
\index{maths}{Ordinal numbers!Accessibility inside epsilon0}
\label{sect:pathes-intro}

@@

Let $s$ be the sequence $\langle s_0,s_1,\dots, s_{l-1} \rangle$. We describe the
existence of such a path with the notation $\alpha\xrightarrow [s]{}\beta$.
We say also that the considered path from $\alpha$ to $\beta$ \emph{starts at [index] $s_0$ and ends at $s_l$}.

@@

Note that, given $\alpha$ and $\beta$, where $\beta < \alpha$, the sequence $s$ which leads from $\alpha$ to $\beta$ is not unique.
Indeed, if $\alpha$ is a limit ordinal, the first element of $s$ can be any integer $i$ such that $\beta<\canonseq{\alpha}{i}$, and if $\alpha$ is a successor ordinal,
then the sequence $s$ can start with any positive integer.

@@

\paragraph*{\gaiasign}
The library~\href{../theories/html/gaia_hydras.GPaths.html}{gaia\_hydras.GPaths} transposes the  notion of path into \gaia's type \texttt{T1}
(please see Sect.\~ref{sect:gpath}).

@@

\vspace{4pt}
\noindent
From~\href{../theories/html/hydras.Hydra.~Hydra_Definitions.html}{Hydra.Hydra\_Definitions}:

@@

@@

The sequence $s$, used to build the sequence of replication factors of the battle, depends on 
$\alpha$ and $\beta$, so we cannot be sure that the generated battle is a genuine standard battle.

The solution to this issue comes  once again from Ketonen and Solovay's article~\cite{KS81}. Instead of considering plain paths, i.e. sequences 
$\alpha_0=\alpha,\alpha_1,\dots,\alpha_k=\beta$ where $\alpha_{j+1}$ is equal
to $\canonseq{\alpha_j}{i_j}$ where $i_j$ is \emph{any} natural number, 
we consider various constraints on these sequences.
In particular, a path is called \emph{standard} if $i_{j+1} = i_j + 1$ for every $j<k$.
It  corresponds to a ``segment'' of some standard battles. 
Please note that the vocabulary on paths is ours, but all the concepts come really from~\cite{KS81}.

@@

\vspace{4pt}
\noindent
\emph{From
  Module~\href{../theories/html/hydras.Epsilon0.Paths.html}{Epsilon0.Paths}}

@@

 \paragraph*{\gaiasign}
 Corollary \texttt{Cor12} and Lemma \texttt{Lemma261} are also available in
\href{../theories/html/gaia_hydras.GPaths.html}{gaia\_hydras.GPaths} 
(please see Sect.\~ref{sect:gpath}).

@@

Our proof follows the proof by Ketonen and Solovay, including its organization as a sequence of lemma.  Since it is a non-trivial proof, we will comment its main steps below   (see Figure~\vref{fig:belle-preuve-1} to Figure~\vref{fig:fin-belle-preuve}).

@@

Applying \texttt{standard\_path\_to\_zero}, $0$ is reachable from $\alpha$ by some standard path.

@@

\subsection{Back to hydras}
\label{sec:standard-battles-cases}
We are now able to complete our proof that there exists no bounded variant for proving the termination of standard hydra battles. This proof can
be consulted in the module 
\href{../theories/html/hydras.Hydra.Epsilon0_Needed_Std.html}{Hydra.Epsilon0\_Needed\_Std}.
Please note that it has the same global structure as in section~\ref{sec:free-battles-case} 
% ICI !
Applying the  lemmas  \texttt{Lemma2\_6\_1} of the module 
\href{../theories/html/hydras.Epsilon0.Paths.html\#Lemma2_6_1}%
{Epsilon0.pathS}   and 
\href{../theories/html/hydras.Epsilon0.Paths.html\#constant_to_standard_path}%
{\texttt{constant\_to\_standard\_path}},
we can convert any inequality $\beta<\alpha<\epsilon_0$ into a standard path from
$\alpha$ to  $\beta$, then into a fragment of a standard battle from 
$\iota(\alpha)$ to $\iota(\beta)$, hence the inequality $m(\iota(\beta))<m(\iota(\alpha))$.
